---
title: "Modeling with a differential equation - IS609 HW-9"
author: "Sharon Morris"
date: "10/20/2017"
output:
  pdf_document: default
  html_document: default
---

## Page 469: #3
The following data were obtained for the growth of a sheep population introduced into a new environment on the island of Tasmania (adapted from J. Davidson, ‘‘On the Growth of the Sheep Population in Tasmania,’’ Trans. R. Soc. S. Australia 62(1938): 342–346).

$$\begin{tabular}{ c | c c c c c c }
    t(year) & 1814 & 1824 & 1834 & 1844 & 1854 & 1864 \\
    \hline
    P(t) & 125 & 275 & 830 & 1200 & 1750 & 1650 \\
\end{tabular}$$

a. Make an estimate of $M$ by graphing $P(t)$.


```{r}
library(ggplot2)
```


```{r}
t <- c(1814, 1824, 1834, 1844, 1854, 1864)
p <- c(125, 275, 830, 1200, 1750, 1650)

ggplot(data.frame(x=t, y=p), aes(x=x, y=y), fill=p) + 
    geom_point(color='steelblue', size=3) + 
    geom_smooth(method = 'lm') +
    theme_bw() +
    xlab("t(Year)") +
    ylab("P(t)") +
    ggtitle("Plot of Sheep Polutlation vs Year")
```
The limimitling population $M\hat = 1850$ -- the values does not exceed 1750


b. Plot ln[$P/(M-P)]$ against $t$. If a logistic curve seems reasonable, estimate $rM$
and $t*$

```{r}
m <- 1850
ln <- log(p/(m-p))

ggplot(data.frame(x=t, y=ln), aes(x=x, y=y), fill=ln) + 
    geom_point(color='steelblue') + 
    geom_smooth(method = 'lm') + 
    xlab("t(Year)") + 
    ylab(("ln")) +
    ggtitle("Ln against t")

```

The above plot shows an approximate linear relationship

####Estimate $rM$ and $t*$

$$ln[\frac {P}{M-P}]$$
```{r}
(rm <- (ln <- lm(log(p/(m-p)) ~ t)$coefficients[[2]]))

```


##  Page 478: #6
Suggest other phenomena for which the model described in the text might be used.

The equation of concentration comes out to be with initial concentration. The model can be used in the following phenomena other than medicine
a. The model can estimate the concentration of radioactive substances used in astronomy which decays over time
b. The model can be used to estimate the age of rocks and the life of trees and plants

## Page 481: #1
a. Using the estimate that $d_b=0.054v^2$, where $0.054$ has dimension $ft.hr^2/mi^2$, show that the constant $k$ in Equation (11.29) has the value 19.9 $ft/sec^2$.
b. Using the data in Table 2.4, plot$d_b$ in ft versus $v^2/2$ in $ft^2$/$sec^2$ to estimate $1/k$ directly

$$d_b=\frac{-v_0^2}{2k}+\frac{v_0^2}{k}=\frac{v_0^2}{2k}$$
$d_b=0.054v^2=\frac{v^2}{2k}$
$k=\frac{1}{2\space X\space 0.054} \approx 9.259\frac{mi^2}{ft.hr^2}$

$9.259\frac{mi^2}{ft.hr^2}$ x $\frac{5280^2}{1 mi^2}$ x $\frac{1hr^2}{3600^2\space s^2}\space \approx19.918\frac{ft}{s^2}$


b. Using the data in Table 2.4, plot $d_b$ in ft versus $v^2/2$ in $ft^2/sec^2$ to estimate $\frac1k$ directly.

```{r}
db <- c(20, 28, 40.5, 52.5, 72, 92.4, 118, 148.5, 182, 220.5, 266, 318, 376)
v <- c(29.33, 36.667, 44, 51.33, 58.667, 66, 73.33, 80.667, 88, 95.333, 102.667, 110.0, 117.3)

ggplot(data.frame(x=v, y=db), aes(x=x, y=y)) + 
geom_smooth(method = 'lm') +
geom_point(size=3) + 
xlab("V(ft per sec)") + 
ylab("db")
```




